Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-01-17 Manuscript/manuscript.Rmd
manuscript edits

@@

In this study we have developed an ensemble framework that aims to address and compare a variety of data irregularities and characteristics that are associated with the prediction of TAD boundaries. We evaluated domain data obtained from four different resolution contact matrices for the GM12878 cell line including 10 kb, 25 kb, 50 kb, and 100 kb. For each resolution, we examined the inclusion of three different predictor types as the feature space for downstream modeling. We coupled each predictor type with two separate variable reduction techniques. Lastly, we performed a series of three different re-sampling techniques aimed at creating balanced classes. Random forest classification algorithms were performed to both compare each combination of the above ensemble, and investigate which functional genomic features were most associated with the formation of TAD boundaries.  Model performance was evaluated based on a total of four metrics which included test accuracy, area under the ROC curve (AUC), F1 score, and Matthew's Correlation Coefficient (MCC).

@@

Here, TP refers to the number of bins correctly identified as containing a TAD boundary (true positives), FP refers to the number of bins incorrectly identified as containing a TAD boundary (false positives), TN refers to the number of bins correctly identified as not containing a TAD boundary (true negatives), and FN refers to the number of bins incorrectly identified as not containing a TAD boundary. Each of these quantities are obtained from the confusion matrix created by validating the model on the testing set.  We also considered the AUC for each model by computing the average trapezoidal approximations for the curve created by the true positive rates (TPR) and false positive rates (FPR).

For the task of identifying the functional genomic elements that were most predictive of the formation of TAD boundaries, we evaluated the variable importances associated with each predictor of the random forest model. For each tree, the prediction accuracy on the out-of-bag portion of the data is recorded. Then the same is done after permuting each predictor variable. The difference between the two accuracies are then averaged over all trees, and normalized by the standard error [@kuhn2012caret].

>>>>

[???include explanation of why included metrics are better]

@@

In this study we have developed an ensemble framework that aims to address and compare a variety of data irregularities and characteristics that are associated with the prediction of TAD boundaries. We evaluated domain data obtained from four different resolution contact matrices for the GM12878 cell line including 10 kb, 25 kb, 50 kb, and 100 kb. For each resolution, we examined the inclusion of three different predictor types as the feature space for downstream modeling. We coupled each predictor type with two separate variable reduction techniques. Lastly, we performed a series of three different re-sampling techniques aimed at creating balanced classes. Random forest classification algorithms were performed to both compare each combination of the above ensemble, and investigate which functional genomic features were most associated with the formation of TAD boundaries.  Model performance was evaluated based on a total of four metrics which included test accuracy, area under the precision-recall curve (AUPRC), F1 score, and Matthew's Correlation Coefficient (MCC).

@@

Here, TP refers to the number of bins correctly identified as containing a TAD boundary (true positives), FP refers to the number of bins incorrectly identified as containing a TAD boundary (false positives), TN refers to the number of bins correctly identified as not containing a TAD boundary (true negatives), and FN refers to the number of bins incorrectly identified as not containing a TAD boundary. Each of these quantities are obtained from the confusion matrix created by validating the model on the testing set. The evaluation metrics listed above are all considered single threshold measures. That is, they are defined for individual thresholds that separated class prediction. Thererfore they are unable to give an overview of the potential range of performance as thresholds vary. A full list of the single threshold measures used (including additional metrics not reported), along with their respective formulas, are provided in Supplementary Table 6. In order to evaluate the trade-off between precision and recall, we also considered a threshold free measure in form of the area under the precision recall curve (AUPRC).

For the task of identifying the functional genomic elements that were most predictive of the formation of TAD boundaries, we evaluated the variable importances associated with each predictor of the random forest model. For each tree, the prediction accuracy on the out-of-bag portion of the data is recorded. Then the same was done after permuting each predictor variable. The difference between the two accuracies were then averaged over all trees, and normalized by the standard error [@kuhn2012caret].
